Quantifying the Importance of Change for Understanding\\ Differences in Personal Culture

### 2024-08-22 write-ups/manuscript/manuscript.tex
changes until examples.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{fontspec} \setmainfont{Times New Roman} \setlength{\parindent}{3em} \usepackage{caption} \usepackage{setspace} \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Quantifying the Importance of Change for Understanding Differences in Personal Culture},
  pdfauthor={Achim Edelmann; Kevin Kiley; Turgut Keskintürk; Isabella Bouklas; Stephen Vaisey},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Quantifying the Importance of Change for Understanding
Differences in Personal Culture}
\author{Achim Edelmann\footnote{médialab, Science Po} \and Kevin
Kiley\footnote{Department of Sociology and Anthropology, North Carolina
  State University} \and Turgut Keskintürk\footnote{Department of
  Sociology, Duke University} \and Isabella
Bouklas\footnote{Department of Sociology, Duke University} \and Stephen
Vaisey\footnote{Department of Sociology, Duke University}}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Work in the sociology of culture has converged on a general conclusion
that while people sometimes make substantial shifts in their personal
culture over time, most elements of personal culture are characterized
by low rates of persistent change during adulthood. Recent developments
have begun to quantify change over time and initial differences, but it
is unclear how to use these components to understand the relative
importance of change, especially when change is modeled non-linearly. To
advance this debate, we introduce a measure for quantifying the
proportion of systematic variance in panel data attributable to
intrapersonal change. Applying this measure to 609 items from seven
surveys in five countries, we find that although intrapersonal change is
common, it does not explain a large share of systematic variance. As an
extension of the theoretical debate, we demonstrate that this measure
provides new insights when comparing social groups, showing that
intrapersonal change is less common among U.S. college graduates than
among those without a college degree. Our findings provide a new
perspective on several important theoretical debates, as well as a tool
to address new questions.
\end{abstract}

\doublespacing

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

Recent work on this topic has converged on a general conclusion that
people do sometimes make substantial shifts in their personal culture
over time but that most elements of personal culture are characterized
by low rates of persistent change during adulthood (Kiley and Vaisey
2020; Lersch 2023). Work leading to this conclusion adopted a
``tournament of models'' approach to adjudicating different theoretical
perspectives (Lersch 2023: p.~228). In this framework, researchers
formalized different models of change over time and relied on model
selection criteria---primarily the Bayesian Information Criterion
(Raftery 1995)---to classify different questions as demonstrating either
``no change'' or ``any change.'' The size of these piles then serves as
the primary evidence for the truth of different arguments (Kiley and
Vaisey 2020; Vaisey and Kiley 2021).

@@

In this paper, we build on the LCAM by outlining a single metric that
combines components of the LCAM to quantify the relative importance of
intrapersonal change and interpersonal difference for explaining
variation in a population for a variable over time. Our approach
isolates the ``systematic'' components of variance (i.e., it removes
random change and measurement error), to provide a single succinct
summary of how much systematic variation can reasonably be attributed to
people making persistent changes over time. This quantity can then be
used as a quick diagnostic to compare across questions, groups, periods,
or samples to isolate where within-person change is more or less useful
for explaining variation. We demonstrate the utility of this quantity in
two ways.

First, we use 609 items from seven panel surveys from five countries to
quantify the proportion of variance that is attributable to systematic
intrapersonal change over the study period rather than to interpersonal
differences at baseline. This allows us to generate an overall picture
of how much of the total systematic variation in personal culture change
tends to explain. These findings largely echo previous findings but
enable clearer statements about the relative importance of change across
questions, rather than binary claims about whether change ``happens'' or
specific quantification of how much change occurs.

Second, to illustrate how this quantity might be used by a researcher
with a specific topical interest, we investigate differences in the
relative importance of intrapersonal change for people with and without
a college degree. We find that change explains a smaller proportion of
systematic variance among college-educated respondents, suggesting that
college crystallizes one's personal culture rather than fostering an
openness to new information. This offers a new empirical basis to
theorize about the role of education in influencing personal culture.

\hypertarget{background}{%
\section{Background}\label{background}}

\hypertarget{stability-and-change-in-personal-culture}{%
\subsection{Stability and Change in Personal
Culture}\label{stability-and-change-in-personal-culture}}

@@

Drawing on these findings, Lersch (2023) challenged the SDM and AUM as a
``needless dichotomy,'' proposing the ``Life Course Adaption Model''
(LCAM) as an alternative. This model draws on the life course
perspective to model personal culture as a different linear
trajectory\footnote{In supplemental materials, Lersch models age using
  splines as well, with results generally showing the same conclusions.}
over the duration of a panel for each respondent. In doing this, Lersch
rectified three shortcomings of the AUM and SDM. First, the AUM, as
described by Kiley and Vaisey, posits that changes follow a Markov
process where responses at time \(t\) are a function of responses at
time \(t-1\) but not earlier time points. However, earlier life
experiences can (directly and indirectly) mold personal culture when
transitioning to new social roles or into new environments, even if
their initial impact is delayed. For example, childhood events might
influence views on family structures later when individuals form their
own families. Therefore, the LCAM considers influences from earlier than
just \(t-1\) on responses at time \(t\).

@@

\hypertarget{quantification-not-adjudication}{%
\subsection{Quantification, Not
Adjudication}\label{quantification-not-adjudication}}

@@

\hypertarget{formalization}{%
\subsection{Formalization}\label{formalization}}

@@

\hypertarget{bootstrapping-for-confidence-intervals}{%
\subsection{Bootstrapping for Confidence
Intervals}\label{bootstrapping-for-confidence-intervals}}

{[}Explain how we do that here{]}

\hypertarget{example-1-cross-question-comparisons}{%
\section{Example 1: Cross-Question
Comparisons}\label{example-1-cross-question-comparisons}}

@@

\hypertarget{data}{%
\subsection{Data}\label{data}}

@@

\begin{center}
<Table 1 about here>
\end{center}

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

@@

\hypertarget{results}{%
\subsection{Results}\label{results}}

@@

\begin{center}
<Figure 1 about here>
\end{center}

@@

\hypertarget{example-2-college-and-change-in-political-culture}{%
\section{Example 2: College and Change in Political
Culture}\label{example-2-college-and-change-in-political-culture}}

@@

\hypertarget{data-1}{%
\subsection{Data}\label{data-1}}

@@

\hypertarget{results-1}{%
\subsection{Results}\label{results-1}}

@@

\begin{center}
<Figure 3 about here>
\end{center}

@@

\hypertarget{discussion-and-conclusion}{%
\section{Discussion and Conclusion}\label{discussion-and-conclusion}}

@@

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

@@

\hypertarget{implications-for-cultural-sociology}{%
\subsection{Implications for Cultural
Sociology}\label{implications-for-cultural-sociology}}

@@

\hypertarget{implications-for-survey-research}{%
\subsection{Implications for Survey
Research}\label{implications-for-survey-research}}

@@

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

@@

\hypertarget{data-note}{%
\section{Data Note}\label{data-note}}

@@

\theendnotes

\singlespace

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-alvarez2002}{}}%
Alvarez, R. Michael, and John Brehm. 2002. \emph{Hard Choices, Easy
Answers: Values, Information, and {American} Public Opinion}. Princeton,
N.J: Princeton University Press.

\leavevmode\vadjust pre{\hypertarget{ref-alwin2007}{}}%
Alwin, Duane F. 2007. \emph{Margins of {Error}: {A} {Study} of
{Reliability} in {Survey} {Measurement}}. Hoboken, N.J.: John Wiley \&
Sons.

\leavevmode\vadjust pre{\hypertarget{ref-alwin1991}{}}%
Alwin, Duane F., and Jon A. Krosnick. 1991.
{``\href{https://www.jstor.org/stable/2781642}{Aging, {Cohorts}, and the
{Stability} of {Sociopolitical} {Orientations} {Over} the {Life}
{Span}}.''} \emph{American Journal of Sociology} 97(1):169--95.

\leavevmode\vadjust pre{\hypertarget{ref-bail2018}{}}%
Bail, Christopher A., Lisa P. Argyle, Taylor W. Brown, John P. Bumpus,
Haohan Chen, M. B. Fallin Hunzaker, Jaemin Lee, Marcus Mann, Friedolin
Merhout, and Alexander Volfovsky. 2018. {``Exposure to Opposing Views on
Social Media Can Increase Political Polarization.''} \emph{Proceedings
of the National Academy of Sciences} 115(37):9216--21. doi:
\href{https://doi.org/10.1073/pnas.1804840115}{10.1073/pnas.1804840115}.

\leavevmode\vadjust pre{\hypertarget{ref-bardi2009}{}}%
Bardi, Anat, Julie Anne Lee, Nadi Hofmann-Towfigh, and Geoffrey Soutar.
2009. {``The Structure of Intraindividual Value Change.''} \emph{Journal
of Personality and Social Psychology} 97(5):913--29.

\leavevmode\vadjust pre{\hypertarget{ref-bourdieu1990}{}}%
Bourdieu, Pierre. 1990. \emph{The Logic of Practice}. Stanford, Calif. :
Stanford University Press, 1990.

\leavevmode\vadjust pre{\hypertarget{ref-boutyline2017}{}}%
Boutyline, Andrei, and Stephen Vaisey. 2017. {``Belief {Network}
{Analysis}: {A} {Relational} {Approach} to {Understanding} the
{Structure} of {Attitudes}.''} \emph{American Journal of Sociology}
122(5):1371--1447. doi:
\href{https://doi.org/10.1086/691274}{10.1086/691274}.

\leavevmode\vadjust pre{\hypertarget{ref-brett2021}{}}%
Brett, Gordon, and Andrew Miles. 2021. {``Who {Thinks} {How}? {Social}
{Patterns} in {Reliance} on {Automatic} and {Deliberate} {Cognition}.''}
\emph{Sociological Science} 8.

\leavevmode\vadjust pre{\hypertarget{ref-brocic2021}{}}%
Broćić, Miloš, and Andrew Miles. 2021. {``College and the {`{Culture}
{War}'}: {Assessing} {Higher} {Education}'s {Influence} on {Moral}
{Attitudes}.''} \emph{American Sociological Review} 00031224211041094.
doi:
\href{https://doi.org/10.1177/00031224211041094}{10.1177/00031224211041094}.

\leavevmode\vadjust pre{\hypertarget{ref-campbell2016}{}}%
Campbell, Colin, and Jonathan Horowitz. 2016. {``Does {College}
{Influence} {Sociopolitical} {Attitudes}?''} \emph{Sociology of
Education} 89(1):40--58. doi:
\href{https://doi.org/10.1177/0038040715617224}{10.1177/0038040715617224}.

\leavevmode\vadjust pre{\hypertarget{ref-christakis2010}{}}%
Christakis, Nicholas, and James Fowler. 2010. \emph{Connected: {The}
{Amazing} {Power} of {Social} {Networks} and {How} {They} {Shape} {Our}
{Lives}}. HarperCollins Publishers.

\leavevmode\vadjust pre{\hypertarget{ref-dellaposta2018}{}}%
DellaPosta, Daniel. 2018. {``Gay {Acquaintanceship} and {Attitudes}
Toward {Homosexuality}: {A} {Conservative} {Test}.''} \emph{Socius}
4:2378023118798959. doi:
\href{https://doi.org/10.1177/2378023118798959}{10.1177/2378023118798959}.

\leavevmode\vadjust pre{\hypertarget{ref-dellaposta2020}{}}%
DellaPosta, Daniel. 2020. {``Pluralistic {Collapse}: {The} {`{Oil}
{Spill}'} {Model} of {Mass} {Opinion} {Polarization}.''} \emph{American
Sociological Review} 85(3):507--36. doi:
\href{https://doi.org/10.1177/0003122420922989}{10.1177/0003122420922989}.

\leavevmode\vadjust pre{\hypertarget{ref-eaton2009}{}}%
Eaton, Asia A., Penny S. Visser, Jon A. Krosnick, and Sowmya Anand.
2009. {``Social {Power} and {Attitude} {Strength} {Over} the {Life}
{Course}.''} \emph{Personality and Social Psychology Bulletin}
35(12):1646--60. doi:
\href{https://doi.org/10.1177/0146167209349114}{10.1177/0146167209349114}.

\leavevmode\vadjust pre{\hypertarget{ref-elder2003}{}}%
Elder, Glen H., Monica Kirkpatrick Johnson, and Robert Crosnoe. 2003.
{``\href{https://doi.org/10.1007/978-0-306-48247-2_1}{The {Emergence}
and {Development} of {Life} {Course} {Theory}}.''} Pp. 3--19 in
\emph{Handbook of the {Life} {Course}}, \emph{Handbooks of {Sociology}
and {Social} {Research}}, edited by J. T. Mortimer and M. J. Shanahan.
Boston, MA: Springer US.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage[]{mathpazo}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{array}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{float}
\usepackage{caption}
\usepackage{amsmath}
\captionsetup{justification=raggedright, singlelinecheck=false}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdfauthor={Achim Edelmann; Kevin Kiley; Turgut Keskintürk; Isabella Bouklas; Stephen Vaisey},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Quantifying the Importance of Change for Understanding\\
Differences in Personal Culture}
\author{Achim Edelmann\footnote{médialab, Science Po; equal authorship.} \and Kevin
Kiley\footnote{Department of Sociology and Anthropology, North Carolina
  State University; equal authorship.} \and Turgut
Keskintürk\footnote{Department of Sociology, Duke University} \and Isabella
Bouklas\footnote{Department of Sociology, Duke University} \and Stephen
Vaisey\footnote{Department of Sociology, Duke University}}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Work in the sociology of culture has converged on a general conclusion
that while people sometimes make substantial shifts in their personal
culture over time, most elements of personal culture are characterized
by low rates of persistent change during adulthood. Recent developments
have begun to quantify change over time and initial differences, but it
is unclear how to use these components to understand the relative
importance of change, especially when change is modeled non-linearly. To
advance this debate, we introduce a measure for quantifying the
proportion of systematic variance in panel data attributable to
intrapersonal change. Applying this measure to 610 items from seven
surveys in five countries, we find that although intrapersonal change is
common, it does not explain a large share of systematic variance. As an
extension of the theoretical debate, we demonstrate that this measure
provides new insights when comparing social groups, showing that
intrapersonal change is less common among U.S. college graduates than
among those without a college degree. Our findings provide a new
perspective on several important theoretical debates, as well as a tool
to address new questions.
\end{abstract}

\section{Introduction}\label{introduction}

@@

Recent work on this topic has converged on a general conclusion that
people do sometimes make substantial shifts in their personal culture
over time but that most elements of personal culture are characterized
by low rates of persistent change during adulthood (Kiley and Vaisey
2020; Lersch 2023). Work leading to this conclusion adopted a
``tournament of models'' approach to adjudicating different theoretical
perspectives (Lersch 2023: p.~228). In this framework, researchers
formalized different models of change over time and relied on model
selection criteria---primarily the Bayesian Information Criterion
(Raftery 1995)---to classify different questions as demonstrating either
``no change'' or ``any change.'' The size of these piles then serves as
the primary evidence for the truth of different arguments (Kiley and
Vaisey 2020; Lersch 2023; Vaisey and Kiley 2021).

@@

In this paper, we build on the LCAM by outlining a single metric that
combines components of the LCAM to quantify the relative importance of
intrapersonal change and interpersonal difference for explaining
variation in a population for a variable over time. Our approach
isolates the ``systematic'' components of variance (i.e., it removes
random change and measurement error), to provide a single summary of how
much systematic variation can reasonably be attributed to people making
persistent changes over time. This quantity can then be used as a quick
diagnostic to compare across questions, groups, periods, or samples to
isolate where within-person change is more or less useful for explaining
variation. We demonstrate the utility of this quantity in two ways.

First, we use 610 items from seven panel surveys from five countries to
quantify the proportion of variance that is attributable to systematic
intrapersonal change over the study period rather than to interpersonal
differences at baseline. This allows us to generate an overall picture
of how much of the total systematic variation in personal culture change
tends to explain. These findings largely echo previous findings but
enable clearer statements about the relative importance of change across
questions, rather than binary claims about whether change ``happens'' or
specific quantification of how much change occurs.

Second, to illustrate how this quantity might be used by a researcher
with a specific topical or theoretical interest, we investigate
differences in the relative importance of intrapersonal change for
people with and without a college degree. We find that change explains a
smaller proportion of systematic variance among college-educated
respondents, suggesting that college crystallizes one's personal culture
rather than fostering an openness to new information. This offers a new
empirical basis to theorize about the role of education in influencing
personal culture.

\section{Background}\label{background}

\subsection{Stability and Change in Personal
Culture}\label{stability-and-change-in-personal-culture}

@@

Drawing on these findings, Lersch (2023) challenged the SDM and AUM as a
``needless dichotomy,'' proposing the ``Life Course Adaption Model''
(LCAM) as an alternative. This model draws on the life course
perspective to model personal culture as a linear trajectory\footnote{In
  supplemental materials, Lersch models age using splines as well, with
  results generally showing the same conclusions.} over the duration of
a panel for each respondent. In doing this, Lersch rectified three
shortcomings of the AUM and SDM. First, the AUM, as described by Kiley
and Vaisey, posits that changes follow a Markov process where responses
at time \(t\) are a function of responses at time \(t-1\) but not
earlier time points. However, earlier life experiences can (directly and
indirectly) mold personal culture when transitioning to new social roles
or into new environments, even if their initial impact is delayed. For
example, childhood events might influence views on family structures
later when individuals form their own families. Therefore, the LCAM
considers influences from earlier than just \(t-1\) on responses at time
\(t\).

@@

\subsection{Quantification, Not
Adjudication}\label{quantification-not-adjudication}

@@

\subsection{Formalization}\label{formalization}

@@

\subsection{Bootstrapping for Quantifying Sampling
Variability}\label{bootstrapping-for-quantifying-sampling-variability}

Of course, the \(\omega\) measure is itself variable---most particularly
due to sampling variability---such that the composition of durable
changers and stable respondents in a particular sample might affect the
systematic components \(V(D)\) and \(V(C)\). It is thus important to
establish confidence intervals to reflect the uncertainty in the
\(\omega\). We achieve this by incorporating nonparametric bootstraps
with case resampling. In doing this, we first resample the respondents
with replacement and assign a new identifier to the set of new
\emph{respondents}, allowing us to change the composition of level-2
units in our data. Second, we estimate the LCAM in this slightly
perturbed sample, and calculate the \(\omega\). We then replicate this
procedure \(N\) times and establish confidence intervals using the
distribution of bootstrapped omega values.

\section{Example 1: Cross-Question
Comparisons}\label{example-1-cross-question-comparisons}

@@

\subsection{Data}\label{data}

@@

\begin{table}[H]
\centering
\caption{\label{tab:unnamed-chunk-1}The Description of the Data Sources}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}[t]{llcc}
\toprule
Country & Survey & Period & Outcomes\\
\midrule
Australia & Household, Income and Labor Dynamics in Australia (HILDA) & 2001-2021 & 30\\
Germany & Socio-Economic Panel Study (SOEP) & 1984-2020 & 122\\
Great Britian & British Household Panel Survey (BHPS) & 1991-2008 & 83\\
Great Britian & Understanding Society/UK Household Longitudinal Study (UKHLS) & 2009-2020 & 66\\
Switzerland & Swiss Household Panel (SHP) & 1999-2019 & 77\\
\addlinespace
United States & General Social Survey (GSS) & 2006-2012 & 183\\
United States & Panel Study of Income Dynamics (PSID) & 1968-2019 & 49\\
\bottomrule
\end{tabular}}
\end{table}

\subsection{Analysis}\label{analysis}

@@

\subsection{Results}\label{results}

@@

\begin{center}\includegraphics[width=1\linewidth]{../../figures/omegas} \end{center}

@@

\section{Example 2: College and Change in Political
Culture}\label{example-2-college-and-change-in-political-culture}

@@

\subsection{Data}\label{data-1}

@@

\subsection{Results}\label{results-1}

@@

\begin{center}\includegraphics[width=62.5in]{../../figures/colleges} \end{center}

@@

\section{Discussion and Conclusion}\label{discussion-and-conclusion}

@@

\subsection{Limitations}\label{limitations}

@@

\subsection{Implications for Cultural
Sociology}\label{implications-for-cultural-sociology}

@@

\subsection{Implications for Survey
Research}\label{implications-for-survey-research}

@@

\subsection{Conclusion}\label{conclusion}

@@

\section{Data Note}\label{data-note}

@@

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{1}
\bibitem[\citeproctext]{ref-alvarez2002}
Alvarez, R. Michael, and John Brehm. 2002. \emph{Hard Choices, Easy
Answers: Values, Information, and {American} Public Opinion}. Princeton,
N.J: Princeton University Press.

\bibitem[\citeproctext]{ref-alwin2007}
Alwin, Duane F. 2007. \emph{Margins of {Error}: {A} {Study} of
{Reliability} in {Survey} {Measurement}}. Hoboken, N.J.: John Wiley \&
Sons.

\bibitem[\citeproctext]{ref-alwin1991}
Alwin, Duane F., and Jon A. Krosnick. 1991.
{``\href{https://www.jstor.org/stable/2781642}{Aging, {Cohorts}, and the
{Stability} of {Sociopolitical} {Orientations} {Over} the {Life}
{Span}}.''} \emph{American Journal of Sociology} 97(1):169--95.

\bibitem[\citeproctext]{ref-bail2018}
Bail, Christopher A., Lisa P. Argyle, Taylor W. Brown, John P. Bumpus,
Haohan Chen, M. B. Fallin Hunzaker, Jaemin Lee, Marcus Mann, Friedolin
Merhout, and Alexander Volfovsky. 2018. {``Exposure to Opposing Views on
Social Media Can Increase Political Polarization.''} \emph{Proceedings
of the National Academy of Sciences} 115(37):9216--21. doi:
\href{https://doi.org/10.1073/pnas.1804840115}{10.1073/pnas.1804840115}.

\bibitem[\citeproctext]{ref-bardi2009}
Bardi, Anat, Julie Anne Lee, Nadi Hofmann-Towfigh, and Geoffrey Soutar.
2009. {``The Structure of Intraindividual Value Change.''} \emph{Journal
of Personality and Social Psychology} 97(5):913--29.

\bibitem[\citeproctext]{ref-bourdieu1990}
Bourdieu, Pierre. 1990. \emph{The Logic of Practice}. Stanford, Calif. :
Stanford University Press, 1990.

\bibitem[\citeproctext]{ref-boutyline2017}
Boutyline, Andrei, and Stephen Vaisey. 2017. {``Belief {Network}
{Analysis}: {A} {Relational} {Approach} to {Understanding} the
{Structure} of {Attitudes}.''} \emph{American Journal of Sociology}
122(5):1371--1447. doi:
\href{https://doi.org/10.1086/691274}{10.1086/691274}.

\bibitem[\citeproctext]{ref-brett2021}
Brett, Gordon, and Andrew Miles. 2021. {``Who {Thinks} {How}? {Social}
{Patterns} in {Reliance} on {Automatic} and {Deliberate} {Cognition}.''}
\emph{Sociological Science} 8.

\bibitem[\citeproctext]{ref-brocic2021}
Broćić, Miloš, and Andrew Miles. 2021. {``College and the {`{Culture}
{War}'}: {Assessing} {Higher} {Education}'s {Influence} on {Moral}
{Attitudes}.''} \emph{American Sociological Review} 00031224211041094.
doi:
\href{https://doi.org/10.1177/00031224211041094}{10.1177/00031224211041094}.

\bibitem[\citeproctext]{ref-campbell2016}
Campbell, Colin, and Jonathan Horowitz. 2016. {``Does {College}
{Influence} {Sociopolitical} {Attitudes}?''} \emph{Sociology of
Education} 89(1):40--58. doi:
\href{https://doi.org/10.1177/0038040715617224}{10.1177/0038040715617224}.

\bibitem[\citeproctext]{ref-christakis2010}
Christakis, Nicholas, and James Fowler. 2010. \emph{Connected: {The}
{Amazing} {Power} of {Social} {Networks} and {How} {They} {Shape} {Our}
{Lives}}. HarperCollins Publishers.

\bibitem[\citeproctext]{ref-dellaposta2018}
DellaPosta, Daniel. 2018. {``Gay {Acquaintanceship} and {Attitudes}
Toward {Homosexuality}: {A} {Conservative} {Test}.''} \emph{Socius}
4:2378023118798959. doi:
\href{https://doi.org/10.1177/2378023118798959}{10.1177/2378023118798959}.

\bibitem[\citeproctext]{ref-dellaposta2020}
DellaPosta, Daniel. 2020. {``Pluralistic {Collapse}: {The} {`{Oil}
{Spill}'} {Model} of {Mass} {Opinion} {Polarization}.''} \emph{American
Sociological Review} 85(3):507--36. doi:
\href{https://doi.org/10.1177/0003122420922989}{10.1177/0003122420922989}.

\bibitem[\citeproctext]{ref-eaton2009}
Eaton, Asia A., Penny S. Visser, Jon A. Krosnick, and Sowmya Anand.
2009. {``Social {Power} and {Attitude} {Strength} {Over} the {Life}
{Course}.''} \emph{Personality and Social Psychology Bulletin}
35(12):1646--60. doi:
\href{https://doi.org/10.1177/0146167209349114}{10.1177/0146167209349114}.

\bibitem[\citeproctext]{ref-elder2003}
Elder, Glen H., Monica Kirkpatrick Johnson, and Robert Crosnoe. 2003.
{``\href{https://doi.org/10.1007/978-0-306-48247-2_1}{The {Emergence}
and {Development} of {Life} {Course} {Theory}}.''} Pp. 3--19 in
\emph{Handbook of the {Life} {Course}}, \emph{Handbooks of {Sociology}
and {Social} {Research}}, edited by J. T. Mortimer and M. J. Shanahan.
Boston, MA: Springer US.

\bibitem[\citeproctext]{ref-feldman1992}
Feldman, Stanley, and John Zaller. 1992. {``The {Political} {Culture} of
{Ambivalence}: {Ideological} {Responses} to the {Welfare} {State}.''}
\emph{American Journal of Political Science} 36(1):268--307. doi:
\href{https://doi.org/10.2307/2111433}{10.2307/2111433}.
